---
title: "Qualitative Visualization"
format: pdf
editor: visual
---

## Qualitative Research Questions

1.  How does genre affect voting average?
2.  How does genre affect runtime?
3.  How does genre affect revenue?
4.  What are the ratings, revenues, and runtimes across 3 popular movie franchises?

## Load Packages

```{r}
#| label: lst-load.pckgs
#| include: false
library(janitor)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(rvest)
library(plotly)
library(esquisse)
```

## Read in, Clean, and Wrangle Data

```{r}
#| label: lst-data
#| include: false
##Importing the Data----
fantasyRaw <- read_csv(
  file = "~/Desktop/STAT184/fantasy.csv"
)

actionRaw <- read_csv(
  file = "~/Desktop/STAT184/action.csv"
)

horrorRaw <- read_csv(
  file = "~/Desktop/STAT184/horror.csv"
)

mysteryRaw <- read_csv(
  file = "~/Desktop/STAT184/mystery.csv"
)

##Merging the Data----
moviesRaw <- full_join(
  x = fantasyRaw,
  y = actionRaw
) %>%
  full_join(
    y = horrorRaw
  ) %>%
  full_join(
    y = mysteryRaw
  ) 

##Cleaning the Data----
moviesCleaned <- moviesRaw %>%
  rename(revenue = `gross(in $)`
  ) %>%
  dplyr:: select(-movie_id, -description, -director_id, -star_id
  ) %>%
  drop_na() %>%
  filter(!grepl('19', year)) %>%
  filter(!duplicated(movie_name)) %>%
  mutate(runtime = readr::parse_number(runtime))

##Listing Only Relevant Movies----
relevantMovies <- moviesCleaned %>%
  separate_wider_delim(
    cols = genre,
    delim = ",",
    names = c("Genre1", "Genre2", "Genre3"),
    too_few = "align_start"
  ) %>%
  pivot_longer(
    cols = starts_with("Genre"),
    names_to = "genreNumber",
    values_to = "genre"
  ) %>%
 mutate(genre = case_match(
   genre,
   " Action" ~ "Action",
   " Mystery" ~ "Mystery",
   " Fantasy" ~ "Fantasy",
  " Horror" ~ "Horror", 
  .default = genre
 )) %>%
  drop_na() %>%
  filter(
    genre == "Action" | 
    genre == "Horror" | 
    genre == "Mystery" | 
    genre == "Fantasy") %>%
  select(-genreNumber)



##Getting Summary Statistics----
info <- list(
  Count = ~as.double(n()),
  Min = ~as.double(min(.x)),
  Q1 = ~as.double(quantile(.x,probs = 0.25, na.rm = TRUE)),
  Median = ~as.double(median(.x)),
  Avg = ~as.double(mean(.x)),
  Q3 = ~as.double(quantile(.x,probs = 0.75, na.rm = TRUE)),
  Max = ~as.double(max(.x))
)

moviesSummary <- relevantMovies %>%
  group_by(genre) %>%
  summarize(across(c(revenue,runtime), info)) %>%
  select(-runtime_Count) %>%
  drop_na() %>%
  rename(count = revenue_Count) 

##Film Franchises----
harryPotterMovies <- relevantMovies %>%
  filter(grepl('Harry Potter', movie_name)) %>%
  select(-star, -genre)

harryPotterSummary <- harryPotterMovies %>%
  summarize(across(c(revenue,runtime), info)) %>%
  select(-runtime_Count) %>%
  drop_na() %>%
  rename(count = revenue_Count) 

piratesMovies <- relevantMovies  %>%
  filter(grepl('Pirates of the Caribbean:', movie_name)) %>%
  select(-star, -genre) %>%
  filter(!duplicated(movie_name))

piratesSummary <- piratesMovies %>%
  summarize(across(c(revenue,runtime), info)) %>%
  select(-runtime_Count) %>%
  drop_na() %>%
  rename(count = revenue_Count) 

spiderMovies <- relevantMovies  %>%
  filter(grepl('Spider-Man', movie_name)) %>%
  select(-star, -genre) %>%
  filter(!duplicated(movie_name))

spiderSummary <- spiderMovies %>%
  summarize(across(c(revenue,runtime), info)) %>%
  select(-runtime_Count) %>%
  drop_na() %>%
  rename(count = revenue_Count)
```

## Genre and Rating

```{r}
#| label: fig-gen.rat
#| echo: false
ggplot(relevantMovies) +
  aes(x = rating, y = genre) +
  geom_boxplot(fill = "#A569FF") +
  labs(
    x = "Average Movie Rating",
    y = "Genres",
    title = "Average Rating by Genre"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 20L,
    face = "bold",
    hjust = 0.5),
    axis.title.y = element_text(size = 15L),
    axis.title.x = element_text(size = 15L),
    axis.text.y = element_text(size = 13L),
    axis.text.x = element_text(size = 13L)
  )
```

## Genre and Revenue

```{r}
#| label: fig-gen.rev
#| echo: false
ggplot(relevantMovies) +
  aes(x = genre, y = revenue) +
  geom_bar(stat = "summary", fun = "sum", fill = "#4682B4") +
  labs(
    x = "Genre",
    y = "Revenue",
    title = "Genre vs Largest Revenue"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 20L,
    face = "bold",
    hjust = 0.5),
    axis.text.y = element_text(size = 13L),
    axis.text.x = element_text(size = 13L)
  )
```

## Genre vs Runtime

```{r}
#| label: fig-gen.run
#| echo: false
ggplot(relevantMovies) +
  aes(x = runtime, y = genre) +
  geom_boxplot(fill = "#65BA65") +
  labs(
    x = "Movie Runtimes",
    y = "Genres",
    title = "Runtimes vs Genres"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 20L,
    face = "bold",
    hjust = 0.5),
    axis.title.y = element_text(size = 15L),
    axis.title.x = element_text(size = 15L)
  )
```

## Harry Potter Movies

```{r}
#| label: fig-potter
#| echo: false
ggplot(harryPotterMovies) +
  aes(
    x = movie_name,
    y = rating,
    fill = revenue,
    colour = runtime
  ) +
  geom_bar(stat = "summary", fun = "sum") +
  scale_fill_gradient() +
  scale_color_distiller(palette = "OrRd") +
  labs(
    x = "Movie Title",
    y = "Average Movie Rating",
    title = "Rating, Revenue, and Runtime of Harry Potter Movies",
    fill = "Revenue",
    color = "Runtime"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 12L,
    face = "bold",
    hjust = 0.5),
    axis.title.y = element_text(size = 10L),
    axis.title.x = element_text(size = 10L),
    axis.text.x = element_text(size = 2L)
  )
```

## Pirates of the Caribbean Movies

```{r}
#| label: fig-pirates
#| echo: false
ggplot(piratesMovies) +
  aes(
    x = movie_name,
    y = rating,
    fill = revenue,
    colour = runtime
  ) +
  geom_bar(stat = "summary", fun = "sum") +
  scale_fill_gradient() +
  scale_color_distiller(palette = "OrRd") +
  labs(
    x = "Movie Title",
    y = "Average Movie Rating",
    title = "Rating, Revenue, and Runtime of Pirates Movies",
    fill = "Revenue",
    color = "Runtime"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 13L,
    face = "bold",
    hjust = 0.5),
    axis.title.y = element_text(size = 10L),
    axis.title.x = element_text(size = 10L),
    axis.text.x = element_text(size = 3L)
  )
```

## Spiderman Movies

```{r}
#| label: fig-spider
#| echo: false
ggplot(spiderMovies) +
  aes(
    x = movie_name,
    y = rating,
    fill = revenue,
    colour = runtime
  ) +
  geom_bar(stat = "summary", fun = "sum") +
  scale_fill_gradient() +
  scale_color_distiller(palette = "OrRd") +
  labs(
    x = "Movie Title",
    y = "Average Movie Rating",
    title = "Rating, Revenue, and Runtime of Spiderman Movies",
    fill = "Revenue",
    color = "Runtime"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 13L,
    face = "bold",
    hjust = 0.5),
    axis.title.y = element_text(size = 10L),
    axis.title.x = element_text(size = 10L),
    axis.text.x = element_text(size = 3L)
  )
```


##Summary Statistics of Movies by Genre

```{r}
#| label: tbl-mov.sum
#| echo: false
moviesSummary %>%
  kable(
    caption = "Summary Statistics of Movies by Genre"
  ) %>%
  kableExtra::kable_classic()
```



