---
title: "Qualitative Visualization"
format: pdf
editor: visual
---

## Qualitative Research Questions

1.  How does genre affect voting average? - Qual
2.  How does genre affect run time? - Qual
3.  How does genre affect revenue? - Qual
4.  What is the relationship between genre, revenue, runtime, and voting average. - Qual

## Install and Load Packages

```{r}
install.packages("ggplot2")
install.packages("janitor")
install.packages("knitr")
install.packages("kableExtra")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("tidyr")
install.packages("dplyr")
install.packages("rvest")

library(janitor)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(rvest)

```

## Read in Data

```{r}
moviesRaw <- read_csv(
  file = "~/Desktop/STAT184/IMDBMovies.csv"
)

moviesCleaned <- moviesRaw %>%
  dplyr:: select(-id, -vote_average, -vote_count, -overview, 
  -backdrop_path, -homepage, -tconst, -poster_path, -tagline, -keywords, 
  -directors, -writers, -cast, -original_title, -popularity
  ) %>%
  filter(adult == FALSE) %>%
  filter(grepl('English', spoken_languages)) %>%
  filter(original_language == "en") %>%
  dplyr:: select(-original_language, -spoken_languages, -adult, -production_companies) %>%
  filter(!grepl('19', release_date)) %>%
  filter(status == "Released") %>%
  filter(revenue > 1) %>%
  filter(runtime >= 30 ) %>%
  filter(budget >=  1000) %>%
  filter(numVotes >=  1000) %>%
  filter(!duplicated(title)) %>%
  filter(!grepl('UFC', title)) %>%
  drop_na() %>%
  separate_wider_delim(
    genres,
    ",",
    names = c("genre1", "genre2", "genre3"),
    too_few = c("align_start"),
    too_many = c("drop")
  )
View(moviesCleaned2)
```

## Summary Stats Table

```{r}
afFreq <- afIndiv %>% tabyl(Rank, Branch, Gender)
afFreqF <- afFreq[["F"]] %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns %>%
  adorn_title("combined", col_name = "Female Armed Forces Members")
kable(afFreqF)


kable(
  diamondsStats,
  col.names = c("Cut", "Count", "Minimum", "Q1", "Q2", "Median", "Q3", "Q4", "Maximum", "Mean", "SD")
  )


##Step 3
penguinModel <- lm(
  formula = body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm + sex + island,
  data = penguinsClean
)

##Step 4
kable(
  penguinModel[["coefficients"]],
  col.names = c("Coefficient", "Value")
)

```

## Genre and Rating

## Genre and Runtime

## Genre and Revenue

## All Variables

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
