---
title: "Quantitative Visualization"
format: pdf
editor: visual
---

## Quantitative Research Questions

1.  What's the relationship between revenue and voting average? - Quant
2.  How does run time affect voting average? - Quant
3.  How does run time affect revenue? - Quant
4.  What is the relationship between genre, revenue, runtime, and voting average. - Quant

## Load Libraries and Packages

```{r}
install.packages("ggplot2")
install.packages("janitor")
install.packages("knitr")
install.packages("kableExtra")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("tidyr")
install.packages("dplyr")
install.packages("rvest")
install.packages("esquisse")


library(janitor)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(tidyr)
library(rvest)
library(plotly)
library(esquisse)
```

## Read and Data Wrangle

```{r}

moviesRaw <- read_csv(
  file = "~/Desktop/STAT184/IMDBMovies.csv"
)

moviesCleaned <- moviesRaw %>%
  dplyr:: select(-id, -vote_average, -vote_count, -overview, 
                 -backdrop_path, -homepage, -tconst, -poster_path, -tagline, -keywords, 
                 -directors, -writers, -cast, -original_title, -popularity
  ) %>%
  filter(adult == FALSE) %>%
  filter(grepl('English', spoken_languages)) %>%
  filter(original_language == "en") %>%
  dplyr:: select(-original_language, -spoken_languages, -adult) %>%
  filter(!grepl('19', release_date)) %>%
  filter(status == "Released") %>%
  filter(revenue > 1) %>%
  filter(runtime >= 30 ) %>%
  filter(runtime <= 220) %>%
  filter(budget >= 1000) %>%
  filter(numVotes >= 1000) %>%
  filter(!duplicated(title)) %>%
  filter(!grepl('UFC', title)) %>%
  drop_na()

```

## Revenue vs. Average Rating

```{r}
ggplot(
  data = moviesCleaned,
  mapping = aes(
    x = averageRating,
    y = revenue
  )
)+
  geom_point(size = 1)+
  labs(
    x = "Average Rating (out of 10)",
    y = "Revenue ($)",
    title = "Revenue vs. Average Rating"
  )+
  theme_bw()+
  theme(
    text = element_text(family="Times New Roman")
  )
```

## Runtime vs. Average Rating

```{r}
ggplot(
  data = moviesCleaned,
  mapping = aes(
    x = averageRating,
    y = runtime
  )
)+
  geom_point(size = 1)+
  labs(
    x = "Average Rating (out of 10)",
    y = "Runtime (mins)",
    title = "Runtime vs. Average Rating"
  )+
  theme_bw()+
  theme(
    text = element_text(family="Times New Roman")
  )
```

## Revenue vs. Runtime

```{r}
ggplot(
  data = moviesCleaned,
  mapping = aes(
    x = runtime,
    y = revenue
  )
)+
  geom_point(size = 1)+
  labs(
    x = "Runtime (mins)",
    y = "Revenue ($)",
    title = "Revenue vs. Runtime"
  )+
  theme_bw()+
  theme(
    text = element_text(family="Times New Roman")
  )
```

## Revenue vs. Runtime with Genre and Average Rating

```{r}
ggplot(
  data = moviesCleaned,
  mapping = aes(
    x = runtime,
    y = revenue,
    z = averageRating,
    color = genres,
    #shape = averageRating
  )
)+
  geom_line(linewidth = 1)+
  labs(
    x = "Runtime (mins)",
    y = "Revenue ($)",
    z = "Average Rating",
    color = "Genre",
    #shape = "Average Rating
    title = "Revenue vs. Runtime with Genre and Average Rating"
  )+
  scale_color_manual(
    values = c("ff3333", "ffa733", "e2ff33", "40ff33", "33ffc1", "3397ff",
               "333bff", "b133ff", "ff33cb", "ff336f", "ffd5b2", "ffb2c9",
               "f7ffb2", "b2ffc0", "b2f9ff", "b4b2ff", "e5b2ff", "ffb2b2",
               "534a4a")
  )
  theme_bw()+
  theme(
    legend.position = "top",
    text = element_text(family="Times New Roman")
  )
```

-   need to edit

The `echo: false` option disables the printing of code (only output is displayed).
